Competing Bandits:\\The Perils of Exploration under Competition

### 2018-11-05 paper/content/barriers.tex
compete

\gaedit{Even though welfare is highest in the temporary monopoly case there are other reasons one may worry about a single firm having a substantial market share. In this section we explore what drives the substantial market share of the temporary monopolist and discuss how this relates to thinking about data as a barrier to entry.}

@@

\gaedit{How does this help us in thinking about data as a barrier to entry? The result that the data advantage is stronger under \TS than \DEG and \DG gives evidence that data quality, and not just quantity, matters in thinking about the competitive advantage of data. In particular, in the incumbency period \TS gathers data that may be hard for the entrant to gather during competition due to its potential reputational consequences (i.e. engaging in exploration), but which are needed in order to learn. If the entrant had some way of acquiring this data without incurring reputational consequences then the data advantage would be substantially weaker. This insight is what differentiates our results from the static viewpoint of the previous literature. While it remains true that, for a given action, increased data on that action has diminishing returns to scale, competitive pressures may make it harder for smaller firms to be able to engage in the exploration necessary to accumulate sufficient data on actions to learn if they are good or bad. Otherwise, they may be more likely to fall into the death-spiral effect discussed earlier. This effect arises when thinking about the issue from the perspective of competition dynamics.}

>>>>

\gaedit{Even though welfare is highest in the temporary monopoly case there are other reasons one may worry about a single firm having a substantial market share. In this section we explore what drives the substantial market share of the temporary monopolist and discuss how this relates to thinking about data as a barrier to entry.}\swcomment{maybe articulate the ``other'' reasons?}

@@

\swedit{Our finding also highlights that the amplification of data
  advantage through the competition dynamics. In particular, in the
  incumbency period \TS gathers data that may be hard for the entrant
  to gather during competition due to its potential reputational
  consequences (by engaging in exploration). This small ``head start''
  goes a long way---the incumbent continues to attract data
  efficiently without sacrificing much of its exploration, while the
  entrant tends to struggle to gather data even with limited
  exploration.
}

\swdelete{\gaedit{How does this help us in thinking about data as a barrier to entry? The result that the data advantage is stronger under \TS than \DEG and \DG gives evidence that data quality, and not just quantity, matters in thinking about the competitive advantage of data. In particular, in the incumbency period \TS gathers data that may be hard for the entrant to gather during competition due to its potential reputational consequences (i.e. engaging in exploration), but which are needed in order to learn. If the entrant had some way of acquiring this data without incurring reputational consequences then the data advantage would be substantially weaker. This insight is what differentiates our results from the static viewpoint of the previous literature. While it remains true that, for a given action, increased data on that action has diminishing returns to scale, competitive pressures may make it harder for smaller firms to be able to engage in the exploration necessary to accumulate sufficient data on actions to learn if they are good or bad. Otherwise, they may be more likely to fall into the death-spiral effect discussed earlier. This effect arises when thinking about the issue from the perspective of competition dynamics.}}

### 2018-11-05 paper/content/conclusion.tex
compete

\gaedit{In this paper we considered a setting where firms simultaneously learn from and compete for users. We showed that competition may not always induce firms to commit to algorithms that engage in exploration resulting in welfare losses for consumers. The primary mechanism behind this was that exploration may have short-term reputational consequences that lead to naive algorithms winning in competition that are bad in the long-run. Allowing one firm to have a head-start, or a first-mover advantage, leads to the first-mover having an incentive to play traditionally ``good" bandit algorithms and resulting in significantly better welfare for consumers. Finally, we attempted to isolate the effect that the data advantage of an incumbent has on its ability to garner a considerable market share. Contrary to previous studies and discussions which do not consider competition dynamics, we find that the data advantage is substantial in generating significant market power. Additionally we show that this effect is strongest when an incumbent performs exploration in the incumbency period and generates a high-quality dataset on actions that may be difficult for an entrant to generate due to the pressure of competition.}

>>>>

\gaedit{In this paper we considered a setting where firms simultaneously learn from and compete for users. We showed that competition may not always induce firms to commit to algorithms that engage in exploration resulting in welfare losses for consumers. The primary mechanism behind this was that exploration may have short-term reputational consequences that lead to naive algorithms winning in competition that are bad in the long-run. Allowing one firm to have a head start, or a first-mover advantage, leads to the first-mover having an incentive to play traditionally ``good" bandit algorithms and resulting in significantly better welfare for consumers. Finally, we attempted to isolate the effect that the data advantage of an incumbent has on its ability to garner a considerable market share. Contrary to previous studies and discussions which do not consider competition dynamics, we find that the data advantage is substantial in generating significant market power. Additionally we show that this effect is strongest when an incumbent \swedit{performs efficient exploration in the incumbency period without the pressure from a competitor.}}
